Formalizing Stack Safety as a Security Property

### 2021-07-09 submission3/main.tex
done with pass

Labeled boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of intended access
rights. Unless otherwise marked, sequential addresses retain the previous domain,
as hinted by their colors. Domains describe the intended security role of the address
that the enforcement mechanism must adhere to. Such a domain
can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
figures. Not shown are global variables and code, which would be labeled \(\outside\).
Domains form part of the {\em context} associated with a machine state.
(We will introduce some other parts of the context as needed.)

@@

\begin{table}[]
\centering
\begin{tabular}{c|c|c}
  Bug & Property Violated & MTTF (ms) \\
  \hline
      {\em LOAD\_NO\_CHECK}  & Confidentiality &  \\
      {\em STORE\_NO\_CHECK} & Integrity &  \\
      {\em HEADER\_NO\_INIT} & Integrity &  \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:bug-table}
\end{table}

The mean-time-to-failure (MTTF) for various bugs can be found in
Table~\ref{tab:bug-table}. All experiments were run in a desktop
machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
RAM.
% \leo{Something like: There is really no baseline to compare this
  %against, but include it here for the future/reasons?}
%
The artificially injected bugs weaken the tag propagation
and checking mechanism.
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by not checking the depth equality (row {\em
  LOAD\_NO\_CHECK}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (row {\em STORE\_NO\_CHECK}).

@@

Observable properties allow us to defer enforcement until a property
violation would become visible. Now we can identify when lazy policies
miss such violations. The Lazy Tagging and Clearing
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Our implementation of this policy, {\em PER\_DEPTH\_TAG}, has an error that we find
in testing. If, at depth \(d\), a callee \(A\) writes into its caller's
frame, then returns, and later another callee \(B\) at depth \(d\) reads the same
address and prints it, \(A\) has violated integrity in a way that later becomes visible,
and \(B\) has likewise violated confidentiality!

@@

We test {\em Lazy Per-activation Tagging and Clearing}, in which
each activation is tagged uniquely, and find that it obeys stepwise
confidentiality and observational integrity. To round out our mutations testing 
we also check an equivalent to {\em LOAD\_NO\_CHECK} and a version where
stores succeed but fail to propagate the PC tag, {\em STORE\_NO\_UPDATE}.

\begin{table}[]
\centering
\begin{tabular}{c|c|c|c}
	Bug & Property Violated & MTTF (ms) & Tests \\
  \hline
	{\em PER\_DEPTH\_TAG} & Observational Integrity &  & \\
	{\em LOAD\_NO\_CHECK}  & Observational Integrity &  & \\
	{\em STORE\_NO\_UPDATE} & Observational Integrity &  & \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:lazy-bug-table}
\end{table}

@@

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of
stack safety begins with \citet{SkorstengaardLocal}, who introduced a calling
convention that uses local capabilities to preserve local state
encapsulation and well-bracketed control flow
using a logical relation to reason about the stack safety of concrete
programs.
% Although the required hardware support is readily available
%in capability machines like Cheri,
%% this technique incurs significant
%% costs, because
%it requires the entire unused part of the stack to be
%cleared whenever a security boundary is crossed.
Their logical
relation captures capability safety without ``externally observable
side-effects (like console output or memory access traces)''
and can be used to reason about individual programs.

StkTokens \citep{SkorstengaardSTKJFP} continues this line of work by
defining a more efficient
calling convention that uses linear capabilities for stack and return
pointers.
Informally, local state encapsulation (LSE) is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow (WBCF) is defined as only allowing
returns from the topmost
frame to the immediately adjacent frame below. Formally, this work improves on
\citet{SkorstengaardLocal} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine.
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
\begin{enumerate}[label=(\roman*)]
\item Most importantly, our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
\citet{SkorstengaardSTKJFP}
statically divide the world into trusted functions, which feature additional
dynamic ``reasonableness'' constraints on their behavior and for which their
properties hold, and adversary functions, which may violate those
constraints but are not guaranteed safety. This static division
is insufficient when we consider buggy, but not inherently adversarial,
code --- one can imagine a function that, when called with the wrong
inputs, behaves unreasonably, but under most circumstances can be
considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
to classify such a function as an adversary, and would not guarantee it
protection. We argue that if individual function activations are well-behaved,
they should be protected.

\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via fully
  abstract overlay
  semantics relies on the definition of a ``stack-safe-by-construction''
  abstract machine that is closely tied to an underlying enforcement
  mechanism and actual machine architecture.
  %the the abstract machine being architecturally similar to
  %      the concrete one (beyond the specific call and return semantics that
  %      distinguish them).
  While the fundamental approach can no doubt be ported to other architectures and enforcement
  schemes, each such port will require a new abstract machine definition whose
  safety-by-construction property must be checked afresh.
  Our properties are more generic: since we abstract over most of the machine
  state as well as the sealing and return conventions, they apply directly
  to a range of architectures, code features, and enforcement mechanisms.

\vspace{.7\smallskipamount}

  Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
  abstract machine to pass arguments on the stack, or to support callee-saves
  registers. It might even be viable to construct a safe-by-construction abstract
  coroutine machine in the same style, but this is hypothetical.
  We have implemented these scenarios with
  simple modifications to the domain model.
\end{enumerate}

>>>>

Labeled boxes indicate an assignment of the address to a
{\em domain}---a label reflecting a particular set of intended access
rights. Unless otherwise marked, sequential addresses retain the previous domain,
as hinted by their colors. Domains are descriptive\bcp{??}; it is up to the enforcement
mechanism to ensure that the system adheres to these permissions. Such a domain
can be \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.) These are abbreviated \(U\) and \(S(\dots)\) in the
figures. Not shown are global variables and code, which would be labeled \(\outside\).
Domains form part of the {\em context} associated with a machine state.
(We will introduce some other parts of the context as needed.)

@@

\sna{I am not sure about the MTTF table. Is it as
  valuable at POPL? Do we need to go into more detail
  about metholodology if we use it?}

%\begin{table}[]
%\centering
%\begin{tabular}{c|c|c}
%  Bug & Property Violated & MTTF (ms) \\
%  \hline
%      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
%      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
%      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
%\end{tabular}
%\vspace*{1em}
%\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
%\vspace*{-2em}
%\label{tab:bug-table}
%\end{table}

%The mean-time-to-failure (MTTF) for various bugs can be found in
%Table~\ref{tab:bug-table}. All experiments were run in a desktop
%machine equipped with i7-8700 CPU @ 3.20GHz with 32GB
%RAM.
% \leo{Something like: There is really no baseline to compare this
  %against, but include it here for the future/reasons?}
%
The artificially injected bugs weaken the tag propagation
and checking mechanism.
%
For example, when loading from a stack location, the correct
micro-policy needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by not checking the depth equality (row {\em
  LOAD\_NO\_CHECK}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (row {\em STORE\_NO\_CHECK}).

@@

Observable properties allow us to defer enforcement until a property
violation would become visible. Now we can identify when lazy policies
miss such violations. The Lazy Tagging and Clearing
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity. \apt{A more detailed example
  would be great here, and should not take too much space.}
\apt{Mention here that this was found by testing.}

@@

We test {\em Lazy Per-activation Tagging and Clearing}, in which
each activation is tagged uniquely, and find that it obeys stepwise
confidentiality and observational integrity. Our testing catches the
violation in the original {\em Lazy Tagging and Clearing}, as well as
an equivalent to {\em LOAD\_NO\_CHECK}, which now also violates observational
integrity.\apt{Move latter to above.}

@@

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of stack safety
%begins with \citet{SkorstengaardLocal}, who introduced a calling
%convention that uses local capabilities to preserve local state
%encapsulation and well-bracketed control flow
%using a logical relation to reason about the stack safety of concrete
%programs.
% Although the required hardware support is readily available
%in capability machines like Cheri,
%% this technique incurs significant
%% costs, because
%it requires the entire unused part of the stack to be
%cleared whenever a security boundary is crossed.
%Their logical
%relation captures capability safety without ``externally observable
%side-effects (like console output or memory access traces)''
%and can be used to reason about individual programs.
is that of \citet{SkorstengaardSTKJFP}. Their 
StkTokens system enforces stack safety on hardware equipped with
memory capabilities, similar to those used in the CHERI system~\citep{Woodruff+14,Chisnall+15}, but
with additional support for linear capabilities.
Stack safety is defined as the conjuncton of local state encapsulation (LSE) and
well-bracketed control flow (WBCF). 
Informally, LSE means that the current activiation can only access its own frame, 
and WBCF means that the topmost frame can only return to 
the immediately adjacent frame below. Formally, these properties are defined
with reference to the semantics of a 
capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine.  They call this approach a ``fully abstract overlay semantics.''
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
\begin{enumerate}[label=(\roman*)]
\item Our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
\citet{SkorstengaardSTKJFP}
statically divide the world into trusted functions, which feature additional
dynamic ``reasonableness'' constraints on their behavior and for which their
properties hold, and adversary functions, which may violate those
constraints but are not guaranteed safety. This static division
is insufficient when we consider buggy, but not inherently adversarial,
code --- one can imagine a function that, when called with the wrong
inputs, behaves unreasonably, but under most circumstances can be
considered reasonable. \citet{SkorstengaardSTKJFP} would be forced
to classify such a function as an adversary, and would not guarantee it
protection.
\apt{But see p. 17 para 2 of the JFP version of their paper, where they explicitly
  dispute this characterization!}
We argue that if individual function activations are well-behaved,
they should be protected.

\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via a fully
  abstract overlay
  semantics relies on the definition of a ``stack-safe-by-construction''
  abstract machine that is closely tied to an underlying enforcement
  mechanism and actual machine architecture.
  %the the abstract machine being architecturally similar to
  %      the concrete one (beyond the specific call and return semantics that
  %      distinguish them).
  While the fundamental approach can no doubt be ported to other architectures and enforcement
  schemes, each such port will require a new abstract machine definition whose
  safety-by-construction property must be checked afresh.
  Our properties are more generic: since we abstract over most of the machine
  state as well as the sealing and return conventions, they apply directly
  to a range of architectures, code features, and enforcement mechanisms.
  Moreover, we have implemented extensions such as passing arguments on the stack,
  describing callee-saves registers, and coroutines by making simple modifications
  to our base domain model.
  %% While it may well be possible to extend 
  %% \citet{SkorstengaardSTKJFP}'s approach to cover these features.  
  %% abstract machine to pass arguments on the stack, or to support callee-saves
  %% registers. It might even be viable to construct a safe-by-construction abstract
  %% coroutine machine in the same style, but this is hypothetical.

\end{enumerate}
